---
title: "Normal Distribution"
subtitle: "Normality testing of student shoe size, height and age"
author: "Niall McGuinness, Jane Doe"
date: "`r format(Sys.time(), '%a, %d %B %y at %X')`"
fontsize: "10pt"
output: 
  html_document:
    toc: true
    toc_title: "Table Of Contents"
    toc_float: true
    toc_depth: 2
    fig_caption: false
    lof: true
    number_sections: true
    code_folding: hide
    theme: paper
---

# Useful Links
* [Testing for normality in R](https://www.programmingr.com/statistics/normality-test/) 
* [How to Perform a Shapiro-Wilk Test in R](https://www.statology.org/shapiro-wilk-test-r/)
* [R Markdown Reference Guide](https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf)
* [Setting up your R markdown YAML Header](https://zsmith27.github.io/rmarkdown_crash-course/lesson-4-yaml-headers.html)
* [Setting document date automatically](https://bookdown.org/yihui/rmarkdown-cookbook/update-date.html)
* [R Markdown Theme Gallery](https://www.datadreaming.org/post/r-markdown-theme-gallery/)
* [An overview of color names in R](https://r-graph-gall
   ery.com/42-colors-names.html)
* [Top 7 Packages for Making Beautiful Tables in R](https://towardsdatascience.com/top-7-packages-for-making-beautiful-tables-in-r-7683d054e541)
* [Beautiful tables in R with gtExtras](https://themockup.blog/posts/2022-06-13-gtextras-cran/)

```{r setup and library load, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
cat('\014')
```


# Load & Inspect Data 

## Load data

```{r Load Data, echo=TRUE}
file_name = "shoe_sizes_1.csv"
student_data <- read.csv(file_name)
```

## Inspect Data Structure

```{r Inspect Head, Tail & Structure, echo=TRUE}
# show structure
str(student_data)

# show first N
head(student_data, 10)
# show last N
tail(student_data, 5)

# show a "prettier" paged table
rmarkdown::paged_table(student_data)

# show a "prettier" paged table but ONLY first 8 rows
head_student_data <- head(student_data, 8)
rmarkdown::paged_table(head_student_data)

# show a "prettier" paged table using kable() with caption and decimal digit max length and alignment = centre
knitr::kable(student_data,
  align = c('c'), # a vector with 'c', 'l', or 'r' values
  digits = 2,
  caption = "A caption for your table here")
```

## Mean, Median & Summary

```{r Mean & Median, echo=TRUE}
shoe_size_mean <- mean(student_data$show_size_eu)
shoe_size_std <- sd(student_data$show_size_eu)

age <- 40
```

The value of age is `r age`

We can access variables from inside code chunks by wrapping the variable in a *mini-chunk* of R as shown on the next line.

Median shoe size statistics were generated (M =`r round(shoe_size_mean, 2)`, SD = `r round(shoe_size_std, 2)`). 

It can be useful to create a summary of the data for quick analysis. A summary of the data is shown below:

```{r Summary, echo=TRUE}
summary(student_data)
#boxplot(student_data$height_cm~student_data$gender)
```

# Three of Four Tests for Normality 

## Method 1: Using Histogram / Density Plot

### Testing Height Distribution

```{r Histogram & Density Plot - Height, echo=TRUE}
hist(student_data$height_cm, 
     main = "Student Height",
     xlab = "Height(cm)",
     ylab = "Frequency",
     breaks = 8)
```

### Testing Shoe Size Distribution

```{r Histogram & Density Plot - Show Size, echo=TRUE}
hist(student_data$show_size_eu, 
     main = "Student Shoe Size",
     xlab = "Size(EU)",
     ylab = "Frequency",
     breaks = 6,
     freq = FALSE)
```

## Method 2: Using QQ Plot (Version 1 - using in-built qqnorm())

```{r QQ Plot - Shoe Size, echo=TRUE}
qqnorm(student_data$show_size_eu,
       pch = 2, frame = FALSE)

qqline(student_data$show_size_eu, 
       col = "darkorchid", lwd = 1)

```

## Method 3: Statistical Test: Shapiro-Wilk

```{r Shapiro-Wilk, echo=TRUE}

#perform shapiro-wilk test
result_sw <- shapiro.test(student_data$show_size_eu)
result_sw
```

A `r result_sw$method` was conducted on the shoe size data. From the output obtained we can not assume normality as the p-value (p=`r round(result_sw$p.value, 4)`) is less than 0.05.









